\textbf{Selective DNA and Protein Isolation from Marine Macrophyte Surfaces}

### 2020-07-08 submission/manuscript.Rmd
Process optimization.

## Results

@@

\newpage
## Experimental procedures
### Sampling
Leaves of *Cymodocea nodosa* were sampled in a *Cymodocea nodosa* meadow in the Bay of Saline (\ang{45;7;5} N, \ang{13;37;20} E) and in a *Cymodocea nodosa* meadow invaded by *Caulerpa cylindracea* in the proximity of the village of Funtana (\ang{45;10;39} N, \ang{13;35;42} E). Thalli of *Caulerpa cylindracea* were sampled in the same *Cymodocea nodosa* invaded meadow in Funtana and on a locality of only *Caulerpa cylindracea* located in the proximity of the invaded meadow. Leaves and thalli were collected on the same day in two contrasting seasons, on 4 December 2017 and 18 June 2018. During spring 2018 the *Cymodocea nodosa* meadow in the Bay of Saline decayed to an extent that no leaves could be retrieved (Najdek *et al.*, unpublished data). Leaves and thalli were collected by diving and transported to the laboratory in containers placed on ice and filled with site seawater. Upon arrival to the laboratory, *Cymodocea nodosa* leaves were cut into sections of 1 -- 2 \si{\cm}, while *Caulerpa cylindracea* thalli were cut into 5 -- 8 \si{\cm} long sections. Leaves and thalli were washed three times with sterile artificial seawater (ASW) to remove loosely attached microbial cells.

@@

### Sequence analysis
```{r}
# Calculation of the error rate based on the ATCC MSA-1002 mock community
error <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.error.summary", na="") %>%
  filter(numparents==1) %>%
  mutate(mismatches_weight=weight*mismatches) %>%
  mutate(total_weight=weight*total)
error <- sum(error$mismatches_weight)/sum(error$total_weight)*100
# Calculation of the number of sequences in the negative control
nc <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary") %>%
  filter(taxon=="Root") %>%
  select(NC_1)
```
Obtained sequences were analyzed on the computer cluster Isabella (University Computing Center, University of Zagreb) using mothur (version 1.43.0) [@Schloss2009] according to the MiSeq Standard Operating Procedure (MiSeq SOP; https://mothur.org/wiki/MiSeq_SOP) [@Kozich2013] and recommendations given from the Riffomonas project to enhance data reproducibility (http://www.riffomonas.org/). For alignment and classification of sequences the SILVA SSU Ref NR 99 database (release 138; http://www.arb-silva.de) was used [@Quast2013; @Yilmaz2014]. Sequences classified as chloroplasts by SILVA were exported and reclassifed using mothur and the RDP (Ribosomal Database Project; http://rdp.cme.msu.edu/) training set (version 16) reference files adapted for mothur [@Cole2014]. In comparison to SILVA, RDP allows a more detailed classification of chloroplast sequences. Pipeline data processing and visualization was done using R (version 3.6.0) [@RCoreTeam2019],  package tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2019; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014]. The detailed analysis procedure including the R Markdown file for this paper are available as a GitHub repository (https://github.com/mkorlevic/Korlevic_SelectiveRemoval_EnvironMicrobiol_2020). Based on the ATCC MSA-1002 mock community included in the analysis a sequencing error rate of `r error` \si{\percent} was determined, which is in line with previously reported values for next-generation sequencing data [@Kozich2013; @Schloss2016]. In addition, the negative control processed together with the samples yielded only `r nc` sequences after sequence quality curation.

@@

## Figure Captions
**\autoref{micro_dec}.** \nameref{micro_dec}

**\autoref{micro_jun}.** \nameref{micro_jun}

@@

\newpage
## Figures
```{r out.width="85%", fig.align="center", fig.cap="Confocal microscope images of \\textit{C. nodosa} and \\textit{C. cylindracea} surfaces from the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) sampled on 4 December 2017 and stained with SYBR Green I. Scale bar in all images is 60 \\si{\\um}.\\label{micro_dec}", fig.pos="ht"}
include_graphics("../results/images/confocal_december.jpg")
```

\newpage
```{r out.width="85%", fig.align="center", fig.cap="Confocal microscope images of \\textit{C. nodosa} and \\textit{C. cylindracea} surfaces from the Bay of Funtana (Mixed and Monospecific Settlements) sampled on 19 June 2018 and stained with SYBR Green I. Scale bar in all images is 60 \\si{\\um}.\\label{micro_jun}", fig.pos="ht"}
include_graphics("../results/images/confocal_june.jpg")
```

>>>>

\newpage
## Results

```{r community}
# Loading input data containing sequence abundances and subsequent input data customization
community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary") %>%
  filter(!str_detect(taxon, "^Eukaryota")) %>%
  filter(taxon!="Root")

# Remove Mitochondria sequences and subtract their number from higher taxonomic levels to which they belong
mitochondria <- filter(community, str_detect(taxon, "^Mitochondria$"))$rankID
community <- mutate_at(community, 5:ncol(community), list(~case_when(
    rankID==str_extract(mitochondria, "(\\d+\\.){4}\\d+") ~ . - .[taxon=="Mitochondria"],
    rankID==str_extract(mitochondria, "(\\d+\\.){3}\\d+") ~ . - .[taxon=="Mitochondria"],
    rankID==str_extract(mitochondria, "(\\d+\\.){2}\\d+") ~ . - .[taxon=="Mitochondria"],
    rankID==str_extract(mitochondria, "(\\d+\\.){1}\\d+") ~ . - .[taxon=="Mitochondria"],
    TRUE ~ .))) %>%
  filter(!str_detect(taxon, "^Mitochondria")) %>%
  select(-ATCC_1, -NC_1)

# Calculating the number of sequences
number_sequences <- filter(community, taxon=="Archaea" | taxon=="Bacteria") %>%
  select(total) %>%
  sum()

# Calculating the number of chloroplast sequences
number_chloroplast_sequences <- filter(community, taxon=="Chloroplast") %>%
  select(total) %>%
  sum()

# Calculating the number of sequences without chloroplast sequences
number_sequences_without_chloroplast <- number_sequences - number_chloroplast_sequences

# Calculating the proprotion of chloroplast sequences
proprotion_chloroplast_sequences <- (number_chloroplast_sequences / number_sequences) * 100

# Calculating the minimal and maximal number of sequences
min_max <- filter(community, taxon=="Archaea" | taxon=="Bacteria") %>%
  summarise_at(6:ncol(.), sum) %>%
  summarise(min=min(.), max=max(.))
```

```{r shared}
# Loading OTU/sample data
shared <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.shared")

# Calculating the number of OTUs before rarefaction
no_otus <- select(shared, starts_with("Otu")) %>%
  colnames() %>%
  length()
```

```{r taxonomy}
# Loading input data containing sequence abundances and subsequent input data customization
community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary") %>%
  filter(!str_detect(taxon, "^Eukaryota")) %>%
  filter(taxon!="Root")

# Remove Mitochondria sequences and subtract their number from higher taxonomic levels to which they belong
mitochondria <- filter(community, str_detect(taxon, "^Mitochondria$"))$rankID
community <- mutate_at(community, 5:ncol(community), list(~case_when(
    rankID==str_extract(mitochondria, "(\\d+\\.){4}\\d+") ~ . - .[taxon=="Mitochondria"],
    rankID==str_extract(mitochondria, "(\\d+\\.){3}\\d+") ~ . - .[taxon=="Mitochondria"],
    rankID==str_extract(mitochondria, "(\\d+\\.){2}\\d+") ~ . - .[taxon=="Mitochondria"],
    rankID==str_extract(mitochondria, "(\\d+\\.){1}\\d+") ~ . - .[taxon=="Mitochondria"],
    TRUE ~ .))) %>%
  filter(!str_detect(taxon, "^Mitochondria")) %>%
  select(-ATCC_1, -NC_1) %>%
  group_by(taxlevel) %>%
  mutate_at(5:ncol(.), list(~. / sum(.) * 100)) %>%
  ungroup()

# Loading metadata of each sample and selection of desired information of each sample
metadata <- read_tsv("data/raw/metadata.csv") %>%
  filter(ID %in% colnames(community[, 6:ncol(community)])) %>%
  select(ID, label) %>%
  deframe()

# Selection of groups for plotting
plot <- filter(community, taxlevel==2 |
                 (taxlevel==4 & str_detect(taxon, "^Chloroplast$")) |
                 (taxlevel==3 & str_detect(rankID, filter(community, str_detect(taxon, "^Proteobacteria$"))[[2]]))) %>%
  filter_at(6:ncol(.), any_vars(. >= 1)) %>%
  mutate_at(5:ncol(.), list(~case_when(taxon=="Cyanobacteria" ~ . - .[taxon=="Chloroplast"], TRUE ~ .))) %>%
  mutate_at(5:ncol(.), list(~case_when(taxon=="Proteobacteria" ~ . - sum(.[taxlevel==3 & str_detect(rankID, filter(community, str_detect(taxon, "^Proteobacteria$"))[[2]])]), TRUE ~ .))) %>%
  mutate(taxon=str_replace(taxon, "Proteobacteria", "Other Proteobacteria")) %>%
  mutate(taxon=str_replace_all(taxon, c("unknown_unclassified"="No_Relative", "unknown"="No_Relative"))) %>%
  filter_at(6:ncol(.), any_vars(. >= 1)) %>%
  bind_rows(summarise_all(., list(~ifelse(is.numeric(.), 100-sum(.), paste("Other"))))) %>%
  rename_at(names(metadata), ~unname(metadata)) %>%
  arrange(taxon %in% "No_Relative") %>%
  mutate(mean=rowMeans(select(., contains(c("SCy", "FCyM", "FCaM", "FCa"))), na.rm=TRUE)) %>%
  mutate(sd=rowSds(as.matrix(select(., contains(c("SCy", "FCyM", "FCaM", "FCa")))), na.rm=TRUE)) %>%
  arrange(desc(mean))
```

```{r chloroplast}
# Loading input data containing sequence abundances and subsequent input data customization
community <- read_tsv("data/mothur/chloroplast.pds.wang.tax.summary") %>%
  filter(taxon!="Root")

# Calculating relative abundaces
community <- group_by(community, taxlevel) %>%
  mutate_at(5:ncol(.), list(~. / sum(.) * 100)) %>%
  ungroup()

# Loading metadata of each sample and selection of desired information of each sample
metadata <- read_tsv("data/raw/metadata.csv") %>%
  filter(ID %in% colnames(community[, 6:ncol(community)])) %>%
  select(ID, label) %>%
  deframe()

# Selection of groups for plotting
chloroplast <- filter(community, taxlevel==5) %>%
  rename_at(names(metadata), ~unname(metadata)) %>%
  mutate(mean=rowMeans(select(., contains(c("SCy", "FCyM", "FCaM", "FCa"))), na.rm=TRUE)) %>%
  mutate(sd=rowSds(as.matrix(select(., contains(c("SCy", "FCyM", "FCaM", "FCa")))), na.rm=TRUE)) %>%
  arrange(desc(mean))
```

To assess the removal efficiency of DNA and protein isolation procedures leaves and thalli were stained with SYBR Green I and examined under a confocal microscope before and after treatments were performed. As it is known that marine macrophytes are harboring more algal epiphytes during Autumn and Winter months  [@Reyes2001], the sampling was performed in two contrasting seasons, in December 2017 when higher abundances of algal epiphytes are expected, and in June 2018, when lower abundances could be expected. In addition, to further test the isolation protocols on different macrophyte species, DNA and proteins were isolated from the surfaces of two macrophytes, the seagrass *C. nodosa* and the macroalga *C. cylindracea*. *C. nodosa* was sampled in a monospecific meadow (Bay of Saline) and *C. cylindracea* invaded meadow (Bay of Funtana), while samples of *C. cylindracea* originated from a monospecific settlement and a *C. cylindracea* invaded *C. nodosa* meadow in the Bay of Funtana. *C. nodosa* June samples from the Bay of Saline were missing as the meadow went to a decline from Spring 2018 (M. Najdek, personal communication). Procedures developed for DNA and protein isolation showed an almost complete removal of the surface community on both, samples collected in December 2017 (\autoref{confocal_december}) and June 2018 (\autoref{confocal_june}). In addition, a similar removal efficiency was observed for communities associated with *C. nodosa* and *C. cylindracea*. Also, no effect of station, settlement or isolation procedure (DNA or protein) on the removal efficiency was observed (Figs. \ref{confocal_december} and \ref{confocal_june}).

To evaluate if the obtained DNA is suitable to determine the microbial community structure an Illumina sequencing of the V4 16S rRNA region was performed. Sequencing yielded a total of `r number_sequences` sequences after quality curation and exclusion of eukaryotic, mitochondrial and no relative sequences. The number of sequences classified as chloroplast was `r number_chloroplast_sequences` (`r proprotion_chloroplast_sequences` \si{\percent})˛ After the exclusion of these sequences the total number of retrieved reads dropped to `r number_sequences_without_chloroplast`, ranging from `r min_max$min` and `r min_max$max` sequences per sample (\autoref{supp-nseq_notus}). Even when the highest sequencing effort was applied the rarefaction curves did not level off that is commonly observed in high-throughput 16S rRNA amplicon sequencing procedures (\autoref{supp-rarefaction}). Sequence clustering at a similarity level of 97 \si{\percent} yielded a total of `r no_otus` different OTUs. Taxonomic classification of reads allowed the determination of the macrophyte associated epiphytic community that was mainly composed of: *Alphaproteobacteria* (`r filter(plot, taxon=="Alphaproteobacteria")$mean` ± `r filter(plot, taxon=="Alphaproteobacteria")$sd` \si{\percent}), *Bacteroidota* (`r filter(plot, taxon=="Bacteroidota")$mean` ± `r filter(plot, taxon=="Bacteroidota")$sd` \si{\percent}), *Gammaproteobacteria* (`r filter(plot, taxon=="Gammaproteobacteria")$mean` ± `r filter(plot, taxon=="Gammaproteobacteria")$sd` \si{\percent}), *Desulfobacterota* (`r filter(plot, taxon=="Desulfobacterota")$mean` ± `r filter(plot, taxon=="Desulfobacterota")$sd` \si{\percent}), *Cyanobacteria* (`r filter(plot, taxon=="Cyanobacteria")$mean` ± `r filter(plot, taxon=="Cyanobacteria")$sd` \si{\percent}) and *Planctomycetota* (`r filter(plot, taxon=="Planctomycetota")$mean` ± `r filter(plot, taxon=="Planctomycetota")$sd` \si{\percent}) (\autoref{community}).

Primers used in this study, as in many other 16S rRNA amplicon procedures, also amplify chloroplast 16S rRNA genes. We observed a high proportion of chloroplast sequences in all analyzed samples (`r filter(plot, taxon=="Chloroplast")$mean` ± `r filter(plot, taxon=="Chloroplast")$sd` \si{\percent}) (\autoref{community}). To determine if chloroplast sequences originate from hosts or eukaryotic epiphytic organisms, we exported SILVA classified chloroplast sequences and reclassified them using the RDP (Ribosomal Database Project) training set that allows for a more detailed chloroplast classification. The largest proportion of sequences was classified as Bacillariophyta (`r filter(chloroplast, taxon=="Bacillariophyta")$mean` ± `r filter(chloroplast, taxon=="Bacillariophyta")$sd` \si{\percent}) indicating that the DNA removal procedure did not coextract larger quantities of host DNA (\autoref{chloroplast}).

@@

\newpage
## Experimental procedures
### Sampling
Leaves of *C. nodosa* were sampled in a *C. nodosa* meadow in the Bay of Saline (\ang{45;7;5} N, \ang{13;37;20} E) and in a *C. nodosa* meadow invaded by *C. cylindracea* in the proximity of the village of Funtana (\ang{45;10;39} N, \ang{13;35;42} E). Thalli of *C. cylindracea* were sampled in the same *C. nodosa* invaded meadow in Funtana and on a locality of only *C. cylindracea* located in the proximity of the invaded meadow. Leaves and thalli were collected on the same day in two contrasting seasons, on 4 December 2017 and 18 June 2018. During spring 2018 the *C. nodosa* meadow in the Bay of Saline decayed to an extent that no leaves could be retrieved (Najdek *et al.*, unpublished data). Leaves and thalli were collected by diving and transported to the laboratory in containers placed on ice and filled with site seawater. Upon arrival to the laboratory, *C. nodosa* leaves were cut into sections of 1 -- 2 \si{\cm}, while *C. cylindracea* thalli were cut into 5 -- 8 \si{\cm} long sections. Leaves and thalli were washed three times with sterile artificial seawater (ASW) to remove loosely attached microbial cells.

@@

### Sequence analysis
```{r}
# Calculation of the error rate based on the ATCC MSA-1002 mock community
error <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.error.summary", na="") %>%
  filter(numparents==1) %>%
  mutate(mismatches_weight=weight*mismatches) %>%
  mutate(total_weight=weight*total)
error <- sum(error$mismatches_weight)/sum(error$total_weight)*100
# Calculation of the number of sequences in the negative control
nc <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary") %>%
  filter(taxon=="Root") %>%
  select(NC_1)
```
Obtained sequences were analyzed on the computer cluster Isabella (University Computing Center, University of Zagreb) using mothur (version 1.43.0) [@Schloss2009] according to the MiSeq Standard Operating Procedure (MiSeq SOP; https://mothur.org/wiki/MiSeq_SOP) [@Kozich2013] and recommendations given from the Riffomonas project to enhance data reproducibility (http://www.riffomonas.org/). For alignment and classification of sequences the SILVA SSU Ref NR 99 database (release 138; http://www.arb-silva.de) was used [@Quast2013; @Yilmaz2014]. Sequences classified as chloroplasts by SILVA were exported and reclassifed using mothur and the RDP (http://rdp.cme.msu.edu/) training set (version 16) reference files adapted for mothur [@Cole2014]. In comparison to SILVA, RDP allows a more detailed classification of chloroplast sequences. Pipeline data processing and visualization was done using R (version 3.6.0) [@RCoreTeam2019],  package tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2019; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014; @Bengtsson2020]. The detailed analysis procedure including the R Markdown file for this paper are available as a GitHub repository (https://github.com/mkorlevic/Korlevic_SelectiveRemoval_EnvironMicrobiol_2020). Based on the ATCC MSA-1002 mock community included in the analysis a sequencing error rate of `r error` \si{\percent} was determined, which is in line with previously reported values for next-generation sequencing data [@Kozich2013; @Schloss2016]. In addition, the negative control processed together with the samples yielded only `r nc` sequences after sequence quality curation.

@@

## Figure Captions
**\autoref{confocal_december}.** \nameref{confocal_december}

**\autoref{confocal_june}.** \nameref{confocal_june}

@@

\newpage
## Figures
```{r out.width="85%", fig.align="center", fig.cap="Confocal microscope images of \\textit{C. nodosa} and \\textit{C. cylindracea} surfaces from the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) sampled on 4 December 2017 and stained with SYBR Green I. Scale bar in all images is 60 \\si{\\um}.\\label{confocal_december}", fig.pos="ht"}
include_graphics("../results/images/confocal_december.jpg")
```

\newpage
```{r out.width="85%", fig.align="center", fig.cap="Confocal microscope images of \\textit{C. nodosa} and \\textit{C. cylindracea} surfaces from the Bay of Funtana (Mixed and Monospecific Settlements) sampled on 19 June 2018 and stained with SYBR Green I. Scale bar in all images is 60 \\si{\\um}.\\label{confocal_june}", fig.pos="ht"}
include_graphics("../results/images/confocal_june.jpg")
```

### 2020-07-08 submission/manuscript.tex
Process optimization.

@@

Leaves of \emph{Cymodocea nodosa} were sampled in a \emph{Cymodocea
nodosa} meadow in the Bay of Saline (\ang{45;7;5} N, \ang{13;37;20} E)
and in a \emph{Cymodocea nodosa} meadow invaded by \emph{Caulerpa
cylindracea} in the proximity of the village of Funtana (\ang{45;10;39}
N, \ang{13;35;42} E). Thalli of \emph{Caulerpa cylindracea} were sampled
in the same \emph{Cymodocea nodosa} invaded meadow in Funtana and on a
locality of only \emph{Caulerpa cylindracea} located in the proximity of
the invaded meadow. Leaves and thalli were collected on the same day in
two contrasting seasons, on 4 December 2017 and 18 June 2018. During
spring 2018 the \emph{Cymodocea nodosa} meadow in the Bay of Saline
decayed to an extent that no leaves could be retrieved (Najdek \emph{et
al.}, unpublished data). Leaves and thalli were collected by diving and
transported to the laboratory in containers placed on ice and filled
with site seawater. Upon arrival to the laboratory, \emph{Cymodocea
nodosa} leaves were cut into sections of 1 -- 2 \si{\cm}, while
\emph{Caulerpa cylindracea} thalli were cut into 5 -- 8 \si{\cm} long
sections. Leaves and thalli were washed three times with sterile
artificial seawater (ASW) to remove loosely attached microbial cells.

@@

Obtained sequences were analyzed on the computer cluster Isabella
(University Computing Center, University of Zagreb) using mothur
(version 1.43.0) (Schloss \emph{et al.}, 2009) according to the MiSeq
Standard Operating Procedure (MiSeq SOP;
\url{https://mothur.org/wiki/MiSeq_SOP}) (Kozich \emph{et al.}, 2013)
and recommendations given from the Riffomonas project to enhance data
reproducibility (\url{http://www.riffomonas.org/}). For alignment and
classification of sequences the SILVA SSU Ref NR 99 database (release
138; \url{http://www.arb-silva.de}) was used (Quast \emph{et al.}, 2013;
Yilmaz \emph{et al.}, 2014). Sequences classified as chloroplasts by
SILVA were exported and reclassifed using mothur and the RDP (Ribosomal
Database Project; \url{http://rdp.cme.msu.edu/}) training set (version
16) reference files adapted for mothur (Cole \emph{et al.}, 2014). In
comparison to SILVA, RDP allows a more detailed classification of
chloroplast sequences. Pipeline data processing and visualization was
done using R (version 3.6.0) (R Core Team, 2019), package tidyverse
(version 1.3.0) (Wickham \emph{et al.}, 2019) and multiple other
packages (Xie, 2014, 2015, 2019b, 2019a, 2020; Neuwirth, 2014; Xie
\emph{et al.}, 2018; Allaire \emph{et al.}, 2019; Zhu, 2019). The
detailed analysis procedure including the R Markdown file for this paper
are available as a GitHub repository
(\url{https://github.com/mkorlevic/Korlevic_SelectiveRemoval_EnvironMicrobiol_2020}).
Based on the ATCC MSA-1002 mock community included in the analysis a
sequencing error rate of 0.009 \si{\percent} was determined, which is in
line with previously reported values for next-generation sequencing data
(Kozich \emph{et al.}, 2013; Schloss \emph{et al.}, 2016). In addition,
the negative control processed together with the samples yielded only 2
sequences after sequence quality curation.

@@

@@

@@

\textbf{\autoref{micro_dec}.} \nameref{micro_dec}

\textbf{\autoref{micro_jun}.} \nameref{micro_jun}

@@

\caption{Confocal microscope images of \textit{C. nodosa} and \textit{C. cylindracea} surfaces from the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) sampled on 4 December 2017 and stained with SYBR Green I. Scale bar in all images is 60 \si{\um}.\label{micro_dec}}\label{fig:unnamed-chunk-2}
\end{figure}

@@

\caption{Confocal microscope images of \textit{C. nodosa} and \textit{C. cylindracea} surfaces from the Bay of Funtana (Mixed and Monospecific Settlements) sampled on 19 June 2018 and stained with SYBR Green I. Scale bar in all images is 60 \si{\um}.\label{micro_jun}}\label{fig:unnamed-chunk-3}
\end{figure}

>>>>

To assess the removal efficiency of DNA and protein isolation procedures
leaves and thalli were stained with SYBR Green I and examined under a
confocal microscope before and after treatments were performed. As it is
known that marine macrophytes are harboring more algal epiphytes during
Autumn and Winter months (Reyes and Sansón, 2001), the sampling was
performed in two contrasting seasons, in December 2017 when higher
abundances of algal epiphytes are expected, and in June 2018, when lower
abundances could be expected. In addition, to further test the isolation
protocols on different macrophyte species, DNA and proteins were
isolated from the surfaces of two macrophytes, the seagrass \emph{C.
nodosa} and the macroalga \emph{C. cylindracea}. \emph{C. nodosa} was
sampled in a monospecific meadow (Bay of Saline) and \emph{C.
cylindracea} invaded meadow (Bay of Funtana), while samples of \emph{C.
cylindracea} originated from a monospecific settlement and a \emph{C.
cylindracea} invaded \emph{C. nodosa} meadow in the Bay of Funtana.
\emph{C. nodosa} June samples from the Bay of Saline were missing as the
meadow went to a decline from Spring 2018 (M. Najdek, personal
communication). Procedures developed for DNA and protein isolation
showed an almost complete removal of the surface community on both,
samples collected in December 2017 (\autoref{confocal_december}) and
June 2018 (\autoref{confocal_june}). In addition, a similar removal
efficiency was observed for communities associated with \emph{C. nodosa}
and \emph{C. cylindracea}. Also, no effect of station, settlement or
isolation procedure (DNA or protein) on the removal efficiency was
observed (Figs. \ref{confocal_december} and \ref{confocal_june}).

To evaluate if the obtained DNA is suitable to determine the microbial
community structure an Illumina sequencing of the V4 16S rRNA region was
performed. Sequencing yielded a total of 336,937 sequences after quality
curation and exclusion of eukaryotic, mitochondrial and no relative
sequences. The number of sequences classified as chloroplast was 97,328
(28.9 \si{\percent})˛ After the exclusion of these sequences the total
number of retrieved reads dropped to 239,609, ranging from 22,587 and
52,958 sequences per sample (\autoref{supp-nseq_notus}). Even when the
highest sequencing effort was applied the rarefaction curves did not
level off that is commonly observed in high-throughput 16S rRNA amplicon
sequencing procedures (\autoref{supp-rarefaction}). Sequence clustering
at a similarity level of 97 \si{\percent} yielded a total of 8,360
different OTUs. Taxonomic classification of reads allowed the
determination of the macrophyte associated epiphytic community that was
mainly composed of: \emph{Alphaproteobacteria} (14.9 ± 3.5
\si{\percent}), \emph{Bacteroidota} (12.5 ± 2.4 \si{\percent}),
\emph{Gammaproteobacteria} (11.6 ± 4.3 \si{\percent}),
\emph{Desulfobacterota} (7.8 ± 7.4 \si{\percent}), \emph{Cyanobacteria}
(6.5 ± 4.7 \si{\percent}) and \emph{Planctomycetota} (2.9 ± 1.7
\si{\percent}) (\autoref{community}).

Primers used in this study, as in many other 16S rRNA amplicon
procedures, also amplify chloroplast 16S rRNA genes. We observed a high
proportion of chloroplast sequences in all analyzed samples (33.4 ± 9.4
\si{\percent}) (\autoref{community}). To determine if chloroplast
sequences originate from hosts or eukaryotic epiphytic organisms, we
exported SILVA classified chloroplast sequences and reclassified them
using the RDP (Ribosomal Database Project) training set that allows for
a more detailed chloroplast classification. The largest proportion of
sequences was classified as Bacillariophyta (89.7 ± 5.7 \si{\percent})
indicating that the DNA removal procedure did not coextract larger
quantities of host DNA (\autoref{chloroplast}).

@@

Leaves of \emph{C. nodosa} were sampled in a \emph{C. nodosa} meadow in
the Bay of Saline (\ang{45;7;5} N, \ang{13;37;20} E) and in a \emph{C.
nodosa} meadow invaded by \emph{C. cylindracea} in the proximity of the
village of Funtana (\ang{45;10;39} N, \ang{13;35;42} E). Thalli of
\emph{C. cylindracea} were sampled in the same \emph{C. nodosa} invaded
meadow in Funtana and on a locality of only \emph{C. cylindracea}
located in the proximity of the invaded meadow. Leaves and thalli were
collected on the same day in two contrasting seasons, on 4 December 2017
and 18 June 2018. During spring 2018 the \emph{C. nodosa} meadow in the
Bay of Saline decayed to an extent that no leaves could be retrieved
(Najdek \emph{et al.}, unpublished data). Leaves and thalli were
collected by diving and transported to the laboratory in containers
placed on ice and filled with site seawater. Upon arrival to the
laboratory, \emph{C. nodosa} leaves were cut into sections of 1 -- 2
\si{\cm}, while \emph{C. cylindracea} thalli were cut into 5 -- 8
\si{\cm} long sections. Leaves and thalli were washed three times with
sterile artificial seawater (ASW) to remove loosely attached microbial
cells.

@@

Obtained sequences were analyzed on the computer cluster Isabella
(University Computing Center, University of Zagreb) using mothur
(version 1.43.0) (Schloss \emph{et al.}, 2009) according to the MiSeq
Standard Operating Procedure (MiSeq SOP;
\url{https://mothur.org/wiki/MiSeq_SOP}) (Kozich \emph{et al.}, 2013)
and recommendations given from the Riffomonas project to enhance data
reproducibility (\url{http://www.riffomonas.org/}). For alignment and
classification of sequences the SILVA SSU Ref NR 99 database (release
138; \url{http://www.arb-silva.de}) was used (Quast \emph{et al.}, 2013;
Yilmaz \emph{et al.}, 2014). Sequences classified as chloroplasts by
SILVA were exported and reclassifed using mothur and the RDP
(\url{http://rdp.cme.msu.edu/}) training set (version 16) reference
files adapted for mothur (Cole \emph{et al.}, 2014). In comparison to
SILVA, RDP allows a more detailed classification of chloroplast
sequences. Pipeline data processing and visualization was done using R
(version 3.6.0) (R Core Team, 2019), package tidyverse (version 1.3.0)
(Wickham \emph{et al.}, 2019) and multiple other packages (Xie, 2014,
2015, 2019b, 2019a, 2020; Neuwirth, 2014; Xie \emph{et al.}, 2018;
Allaire \emph{et al.}, 2019; Zhu, 2019; Bengtsson, 2020). The detailed
analysis procedure including the R Markdown file for this paper are
available as a GitHub repository
(\url{https://github.com/mkorlevic/Korlevic_SelectiveRemoval_EnvironMicrobiol_2020}).
Based on the ATCC MSA-1002 mock community included in the analysis a
sequencing error rate of 0.009 \si{\percent} was determined, which is in
line with previously reported values for next-generation sequencing data
(Kozich \emph{et al.}, 2013; Schloss \emph{et al.}, 2016). In addition,
the negative control processed together with the samples yielded only 2
sequences after sequence quality curation.

@@

\leavevmode\hypertarget{ref-Bengtsson2020}{}%
Bengtsson, H. (2020) MatrixStats: Functions that apply to rows and
columns of matrices (and to vectors).

@@

\leavevmode\hypertarget{ref-Reyes2001}{}%
Reyes, J. and Sansón, M. (2001) Biomass and production of the epiphytes
on the leaves of \emph{Cymodocea nodosa} in the Canary Islands.
\emph{Bot Mar} \textbf{44}: 307--313.

@@

\textbf{\autoref{confocal_december}.} \nameref{confocal_december}

\textbf{\autoref{confocal_june}.} \nameref{confocal_june}

@@

\caption{Confocal microscope images of \textit{C. nodosa} and \textit{C. cylindracea} surfaces from the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) sampled on 4 December 2017 and stained with SYBR Green I. Scale bar in all images is 60 \si{\um}.\label{confocal_december}}\label{fig:unnamed-chunk-2}
\end{figure}

@@

\caption{Confocal microscope images of \textit{C. nodosa} and \textit{C. cylindracea} surfaces from the Bay of Funtana (Mixed and Monospecific Settlements) sampled on 19 June 2018 and stained with SYBR Green I. Scale bar in all images is 60 \si{\um}.\label{confocal_june}}\label{fig:unnamed-chunk-3}
\end{figure}
